---
title: "July High DAU Invesitgation"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---

This report analyze the reason of July high DAU by looking new/current/returning users and their retention rate and review our current social decaying model to see if we missed anything in the model or this is an accidental event.

Methodology:
Group A: 1/1/2020-3/8/2020 (Before Covid19)
Group B: 3/9/2020-6/13/2020 (Initial Social Distancing Data)
Group C: 7/1/2020-7/18/2020 (July Data)

The final report can be found [here](https://confluenator.kingsisle.com:8443/display/WIZ/Investigation+of+July+2020+High+User+Volume).

DAU Plot
```{r}
#Packages
library(dplyr)
library(ggplot2)
library(sqldf)
data<-read.csv('./July High DAU Investigation Data.csv')
data$date<-as.Date(data$date,"%m/%d/%Y")
#Update model version date
first_model_date<-as.Date("2020-06-13")
second_model_date<-as.Date("2020-07-18")
#This is exactly data in the latest version
all_model_data<-data%>%
  mutate(actual=ifelse(date>=as.Date("2020-01-01")&date<=second_model_date,actual,NA),
         current=ifelse(date>=as.Date("2020-01-01")&date<=second_model_date,current,NA),
         new=ifelse(date>=as.Date("2020-01-01")&date<=second_model_date,new,NA),
         returning=ifelse(date>=as.Date("2020-01-01")&date<=second_model_date,returning,NA)
)
ggplot(all_model_data,aes(x=date))+
  geom_line(aes(y=actual),color='darkred')+
  geom_line(aes(y=social_distancing),color='steelblue')+
  ggtitle("Acutual DAU", subtitle=paste("From 1/1/2020 to ",second_model_date,sep=''))+
  xlab("Date")+ylab("DAU")+ylim(0,75000)+
  theme_light()
```

There is an increase starting from early July

```{r}
group_a<-all_model_data%>%
  filter(date<='2020-03-08')%>%
  arrange(date)
group_b<-all_model_data%>%
  filter(date>='2020-03-09',date<=first_model_date)%>%
  arrange(date)
group_c<-all_model_data%>%
  filter(date>='2020-07-01',date<=second_model_date)%>%
  arrange(date)
```
User Retention Rate T Test: Group B vs Group C
When P value is lower than 0.05, two groups retentions are significantly different
```{r}
t.test(group_b$current_r,group_c$current_r)
t.test(group_b$new_r,group_c$new_r)
t.test(group_b$returning_r,group_c$returning_r)
```
Current users and new users retention rate in July is significantly higher than March to June
Retention Plot
```{r}
#Plot Retention
ggplot(data=all_model_data%>%filter(!is.na(actual)))+
  geom_line(aes(x=date,y=current_r),color='sky blue',size=1)+
  geom_line(aes(x=date,y=new_r),color='orange',size=1)+
  geom_line(aes(x=date,y=returning_r),color='yellow',size=1)+
  ggtitle("Acutal Retention Rate") + xlab("Date") + ylab("Retention Rate")+
  geom_text(aes(x=as.Date('2020-06-10'), y = 0.9, label = "Current Retention"))+
  geom_text(aes(x=as.Date('2020-06-10'), y = 0.6, label = "Returning Retention"))+
  geom_text(aes(x=as.Date('2020-06-10'), y = 0.4, label = "New Retention"))+
  ylim(0,1)+
  theme_light()
```
```{r}
first_model_data<-data%>%
  mutate(actual=ifelse(date>=as.Date("2020-01-01")&date<=first_model_date,actual,NA),
         current=ifelse(date>=as.Date("2020-01-01")&date<=first_model_date,current,NA),
         new=ifelse(date>=as.Date("2020-01-01")&date<=first_model_date,new,NA),
         returning=ifelse(date>=as.Date("2020-01-01")&date<=first_model_date,returning,NA))%>%
  arrange(date)
first_model_na<-first_model_data%>%filter(!is.na(actual))
cr<-mean(first_model_na$current_r)
nr<-mean(first_model_na$new_r)
rr<-mean(first_model_na$returning_r)
cr_july<-mean(group_c$current_r)
nr_july<-mean(group_c$new_r)
rr_july<-mean(group_c$returning_r)
```
```{r}
result_1<-first_model_data %>%
  mutate(current_new_1=ifelse(date<=first_model_date,current,lag(current)*cr+lag(new)*nr+lag(returning)*rr),
         new_new_1=ifelse(date<=first_model_date,new,lag(new)-10),
         return_new_1=ifelse(date<=first_model_date,returning,lag(returning)-12)
        )
start_row<-nrow(first_model_data%>%filter(!is.na(actual)))+2
end_row<-nrow(first_model_data)
for (i in start_row:end_row){
  result_1[i,15]=result_1[i-1,15]-10
  result_1[i,16]=result_1[i-1,16]-12
  result_2<-result_1
}
result_3<-result_2%>%
  mutate(new_new_1=ifelse(new_new_1>=new_linear,new_new_1,new_linear),
         return_new_1=ifelse(return_new_1>=return_log,return_new_1,return_log),
         current_new_2=current_new_1)
for (i in start_row:end_row){
  result_3[i,14]=result_3[i-1,14]*cr+result_3[i-1,15]*nr+result_3[i-1,16]*rr
  result_3[i,17]=result_3[i-1,17]*cr_july+result_3[i-1,15]*nr_july+result_3[i-1,16]*rr_july
  result_4<-result_3
}
#Current_new_1 is based on Covid19 first model high retention rate and Current_new_2 is based on July high retention rate
```
```{r}
result_5<-result_4%>%
  mutate(actual_dau=current+new+returning,
         december_model_dau=current_log+new_linear+return_log,
         june_model_dau=current_new_1+new_new_1+return_new_1,
         july_model_dau=current_new_2+new_new_1+return_new_1)%>%
  arrange(date)%>%
  select(actual_dau,december_model_dau,june_model_dau,july_model_dau)
final_result<-cbind(data,result_5)
ggplot(final_result,aes(x=date))+
  geom_line(aes(y=actual),color='darkred')+
  geom_line(aes(y=december_model_dau),color='steelblue')+
  geom_line(aes(y=june_model_dau),color='yellow')+
  geom_line(aes(y=july_model_dau),color='green')+
  ggtitle("DAU Models", subtitle="From 1/1/2020 to ")+
  xlab("Date")+ylab("DAU")+ylim(0,75000)+
  theme_light()
write.csv(final_result,'./final_result_1.csv')
```
In July, current users and new users retention rate went up to push DAU up then DAU follows this high retention decaying line till the end of August.
```{r}
covid<-read.csv('./covid.csv')
covid$date<-as.Date(covid$date,"%m/%d/%Y")
data_covid<-sqldf(
  '
  select data.date,data.current_r,data.new_r,daily_case
  from data
  left join covid on data.date=covid.date
  '
  )
cor.test(data_covid$daily_case,data_covid$current_r)
cor.test(data_covid$daily_case,data_covid$new_r)
```
There is a small correlation (>0.2) between current users/new users retention rate and daily Covid 19 cases